---
title: "Hearth Disease logistic Regression"
author: "Daniel Chavarría"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Let's read the data

This comes from the heart disease dataset and is available at the url below, 

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(broom)

raw_data <- "http://archive.ics.uci.edu/ml/machine-learning-databases/heart-disease/processed.cleveland.data" %>% 
  read_csv(col_names = FALSE)

names(raw_data) = c(
  "age",
  "sex",# 0 = female, 1 = male
  "cp", # chest pain
  # 1 = typical angina,
  # 2 = atypical angina,
  # 3 = non-anginal pain,
  # 4 = asymptomatic
  "trestbps", # resting blood pressure (in mm Hg)
  "chol", # serum cholestoral in mg/dl
  "fbs",  # fasting blood sugar if less than 120 mg/dl, 1 = TRUE, 0 = FALSE
  "restecg", # resting electrocardiographic results
  # 1 = normal
  # 2 = having ST-T wave abnormality
  # 3 = showing probable or definite left ventricular hypertrophy
  "thalach", # maximum heart rate achieved
  "exang",   # exercise induced angina, 1 = yes, 0 = no
  "oldpeak", # ST depression induced by exercise relative to rest
  "slope", # the slope of the peak exercise ST segment
  # 1 = upsloping
  # 2 = flat
  # 3 = downsloping
  "ca", # number of major vessels (0-3) colored by fluoroscopy
  "thal", # this is short of thalium heart scan
  # 3 = normal (no cold spots)
  # 6 = fixed defect (cold spots during rest and exercise)
  # 7 = reversible defect (when cold spots only appear during exercise)
  "hd" # (the predicted attribute) - diagnosis of heart disease
  # 0 if less than or equal to 50% diameter narrowing
  # 1 if greater than 50% diameter narrowing
)
```

## Cleaning up the data

Now, we notice that some columns need to be a factor, let's get done with that

```{r, warning = FALSE}

head(raw_data)

heart_disease <- raw_data %>% 
  mutate(
    sex = case_when(sex == 1 ~ "Male",
                    sex == 0 ~ "Female"),
    sex = as_factor(sex),
    cp = as_factor(cp),
    fbs = as_factor(fbs),
    restecg = as_factor(restecg),
    exang = as_factor(exang), 
    slope = as_factor(slope),
    ca = as.integer(ca) %>% as_factor(),
    thal = as.integer(thal) %>% as_factor(),
    hd = case_when(hd == 0 ~ "Healthy",
                   hd >= 1 ~ "Unhealthy") %>% as_factor()
    
         ) %>%
  na.omit() #There are only six rows with missing values

head(heart_disease)

```

Now, every variable is way clearer and we can start doing some processing. Let's first look whether or not the levels make up for
significant amounts of heart disease cases
## Tabular Analysis
```{r}
xtabs(~hd + sex, data = heart_disease)
xtabs(~hd + cp, data = heart_disease)
xtabs(~hd + fbs, data = heart_disease)
xtabs(~hd + restecg, data = heart_disease) #level 1 is worrysome
xtabs(~hd + exang, data = heart_disease)
xtabs(~hd + slope, data = heart_disease)
xtabs(~hd + ca, data = heart_disease)
xtabs(~hd + thal, data = heart_disease)
```
#Let's do now some logistic regression only with the sex

```{r}
xtabs(~hd + sex, data = heart_disease) #There is a clear trend that sex may be a good predictor of heart disease

logistic_sex <- glm(hd ~ sex, data = heart_disease, family = "binomial")

#Let's fist take a look at the coefficients
tidy(logistic_sex)
#The intercept is the log(odds) a Male will be unhealthy That is equal to log(112/89)
male_log_odds <- log(112/89)
#The second coefficient means that the odds of being unhealthy are, on a log scale, -1.27 times lower than if
#a sample has sex=M, that is:
female_log_odds_rate <- log((25/71) / (112/89))
#Using broom's augment, we get the predicted data for the model in the column .fitted
augment(logistic_sex) %>% 
  ggplot(aes(x = sex, y = .fitted, color = sex)) +
  geom_point()

xtabs(~ .fitted + sex, data=augment(logistic_sex))
```

```{r}
logistic <- glm(hd ~ ., data=heart_disease, family="binomial")
tidy(logistic)
 
## Now calculate the overall "Pseudo R-squared" and its p-value
ll_null <- logistic$null.deviance/-2
ll_proposed <- logistic$deviance/-2
 
## McFadden's Pseudo R^2 = [ LL(Null) - LL(Proposed) ] / LL(Null)
(ll_null - ll_proposed) / ll_null
 
## The p-value for the R^2
1 - pchisq(2*(ll_proposed - ll_null), df = (length(logistic$coefficients)-1))

glance(logistic)

## now we can plot the data
augment(logistic, type.predict = "response") %>% 
  arrange(.fitted) %>% 
  rownames_to_column() %>%
  mutate(rowname = as.integer(rowname)) %>%
ggplot(aes(x=rowname, y=.fitted)) +
  theme_light() + 
  geom_point(aes(color=hd), alpha=0.8, shape=4, stroke=2) +
  xlab("Index") +
  ylab("Predicted probability of getting heart disease")
 
ggsave("heart_disease_probabilities.pdf")

```

